---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<b><span class="h1">Binary Logistics</span></b>

<span class="h5">
<b>Modern Social Analysis, HSE SPb</b><br>
*Course:* <b>Quantitative Data Analysis</b>
</span>

<span class="h4"><b>
Vasily Khodakovsky</span>
</b></div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

# Data


```{r include = FALSE}
load("C:/Users/Vasily/Downloads/Logit-Probit.RData")
# Select the desired variables from the "france" dataframe
france <- france[, c("books1", "cinema1", "age", "gender", "edu", "bills", "commun")]


```


```{r}
head(france)
```

<br>
<br>
<br>

### Variables

`books1` - whether the respondent read a book in the last 12 months (no, yes);

`cinema1` - whether the respondent attended cinema in the last 12 months (no, yes);

`age`  - age of the respondent in years;

`gender` - gender of the respondent (male, female);

`edu` - age when the respondent completed education; 

`bills` - whether the respondent had difficulties paying bills in the last 12 months (no, yes);

`commun` - area of residence (rural ares, small/middle town, large town).

<br>
<br>
<br>
<br>
<br>



# Logit

```{r}
library(sjPlot)
model1 = glm(books1 ~ gender + age + edu + bills + commun, family = binomial, data = france)
tab_model(model1) #log of odds ratio
```

<br>

`age` - the log of odds ratio of reading a book over not reading a book increases by 0.02 with each year. With age, the probability of reading a book increases.

`gender` - the log of odds ratio is greater by 0.37 for women compared to men. Women are more likely to read a book.

Economic difficulties and area of residence do not affect the probability of reading books.

<br>
<br>
<br>

```{r}
exp(coef(model1)) #odds ratio
```

<br>

`age` - the odds of reading a book increase by 2% with each year of age (odds - 1).

`gender` - the odds of reading a book are 1.5 higher for women compared to men.

<br>
<br>
<br>

### Model fit; R2

```{r}
library(DescTools)
PseudoR2(model1) #should be between 0.2 and 0.4 for perfect fit
```

<br>
<br>
<br>

### Correctly predicted values

```{r}
library(pscl)
hitmiss(model1) #should be >50%
```

<br>
<br>
<br>


# Probit

```{r}
model2 = glm(books1 ~ gender + age + edu + bills + commun, 
             family = binomial(link = probit), 
             data = france)
tab_model(model2) #z-scores
```

<br>
<br>
<br>

```{r}
library(margins)
summary(margins(model2))
```

<br>
<br>
<br>

```{r}
summary(margins(model1))
```

<br>

On average the probability of reading a book increases by 0.004 with each year of age.

On average for women the probability is greater by 0.07.

<br>
<br>
<br>

```{r}
PseudoR2(model2)
```

<br>

```{r}
hitmiss(model2)
```

