---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<span class="h1">Binary Logistics</span>

<span class="h5">
Modern Social Analysis, HSE SPb<br>
*Course:* Quantitative Data Analysis
</span>

<span class="h4">
Vasily Khodakovsky</span>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

# Data


```{r include = FALSE}
load("C:/Users/Vasily/Downloads/Logit-Probit.RData")
france <- france[, c("books1", "cinema1", "age", "gender", "edu", "bills", "commun")]
```


```{r}
head(france)
```

<br>
<br>
<br>

### Variables

`books1` - whether the respondent read a book in the last 12 months (no, yes);

`cinema1` - whether the respondent attended cinema in the last 12 months (no, yes);

`age`  - age of the respondent in years;

`gender` - gender of the respondent (male, female);

`edu` - age when the respondent completed education; 

`bills` - whether the respondent had difficulties paying bills in the last 12 months (no, yes);

`commun` - area of residence (rural ares, small/middle town, large town).

<br>
<br>
<br>
<br>
<br>



# Logit

```{r}
library(sjPlot)
model1 = glm(books1 ~ gender + age + edu + bills + commun, family = binomial, data = france)
tab_model(model1)
```

<br>

Log of odds ratio

<br>
<br>
<br>

```{r}
exp(coef(model1))
```

<br>

Odds ratio

<br>
<br>
<br>

### Model fit; R2

```{r}
library(DescTools)
PseudoR2(model1)
```

<br>

Should be between 0.2 and 0.4 for perfect fit

<br>
<br>
<br>

### Correctly predicted values

```{r}
library(pscl)
hitmiss(model1)
```

<br>

Should be 50%+

<br>
<br>
<br>


# Probit

```{r}
model2 = glm(books1 ~ gender + age + edu + bills + commun, 
             family = binomial(link = probit), 
             data = france)
tab_model(model2)
```

<br>

Z-scores

<br>
<br>
<br>

```{r}
library(margins)
summary(margins(model2))
```

<br>
<br>
<br>

```{r}
summary(margins(model1))
```

<br>
<br>
<br>

```{r}
PseudoR2(model2)
```

<br>

```{r}
hitmiss(model2)
```

