---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<span class="h1">Causal Inference</span>

<span class="h5">
Modern Social Analysis, HSE SPb<br>
*Course:* Quantitative Data Analysis
</span>

<span class="h4">
Vasily Khodakovsky</span>
</div>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

```


```{r, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

# Packages

```{r}
library(lpdensity)
library(rddensity)
library(rdrobust)
library(foreign)
library(sjPlot)
```

<br>
<br>
<br>
<br>
<br>

# Data

```{r, include=FALSE}
data = read.dta("C:/Users/Vasily/Downloads/CIT_2018_Cambridge_polecon.dta")
```

<br>

```{r}
head(data)
```

<br>
<br>
<br>
<br>
<br>

# Analysis

<br>

### Regression discontinuity design

```{r}
rdplot(data$Y, data$X, nbins = c(2500, 500), p = 0, col.lines = "red", 
       col.dots = "black", title = "", 
       x.label = "Islamic Margin of Victory", 
       y.label = "Female High School Percentage", 
       y.lim = c(0,70))
```

<br>
<br>
<br>

### Checking the difference in the N of observations in groups

```{r}
summary(rdrobust(data$lpop1994, data$X))
```

<br>
<br>
<br>

### Checking the difference in the density around cutting point

```{r}
summary(rddensity(data$X))
```

<br>
<br>
<br>

### Checking the fake cut-off

```{r}
summary(rdrobust(data$Y[data$X >= 0], data$X[data$X >= 0], c = 1))
```

<br>
<br>
<br>

### Setting the number of intervals (30 intervals from both sides)

<br>

##### Linear effect:

```{r}
rdplot(data$Y, data$X, nbins = c(30,30), p = 1,
       title = "Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.2",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       binselect = 'esmv', y.lim = c(0,25))
```

<br>
<br>
<br>

##### Quadratic effect:

```{r}
rdplot(data$Y, data$X, nbins = c(30,30), p = 2,
       title = "Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.2",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       binselect = 'esmv', y.lim = c(0,25))
```

<br>
<br>
<br>

##### 3rd degree polynomial:

```{r}
rdplot(data$Y, data$X, nbins = c(30,30), p = 3,
       title = "Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.2",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       binselect = 'esmv', y.lim = c(0,25))
```

<br>
<br>
<br>

##### Intervals are selected automatically:

```{r}
rdplot(data$Y, data$X,  binselect = 'qs', p = 1,
       title = "RD plot with IMSE-optimal quantile-spaced binning 
       Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.7",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       x.lim = c(-100,100), y.lim = c(0,25))
```

<br>
<br>
<br>

```{r}
rdplot(data$Y, data$X,  binselect = 'qs', p = 2,
       title = "RD plot with IMSE-optimal quantile-spaced binning
       Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.7",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       x.lim = c(-100,100), y.lim = c(0,25))
```

<br>
<br>
<br>

```{r}
rdplot(data$Y, data$X,  binselect = 'qs', p = 3,
       title = "RD plot with IMSE-optimal quantile-spaced binning
       Cattaneo, Idrobo, and Titiunik 2017a, Fig. 3.7",
       x.label = "Islamic Margin of Victory",
       y.label = "Female High School Percentage", 
       x.lim = c(-100,100), y.lim = c(0,25))
```

<br>
<br>
<bt>

```{r}
summary(rdrobust(data$Y, data$X, kernel = 'triangular', 
                 c = 0, p = 2, bwselect = 'mserd'))
```

The robust effect is non-significant.

<br>
<br>
<br>

### Including controls

```{r}
controls = cbind(data$vshr_islam1994, data$partycount, data$lpop1994,
                 data$merkezi, data$merkezp, data$subbuyuk, data$buyuk)
```

<br>

```{r}
colnamesZ = c("vshr_islam1994", "partycount", "lpop1994", 
              "merkezi", "merkezp", "subbuyuk", 
              "buyuk") #doesn't work well with categorical variables
```

<br>

```{r}
summary(rdrobust(data$Y, data$X, covs = controls, kernel = 'triangular', 
                 p = 2, bwselect = 'mserd'))
```

The effect is non-significant.

<br>
<br>
<br>
<br>
<br>

# Instrumental variables 

```{r}
library(AER)
library(npsf)
```

<br>

```{r}
data(mroz)
head(mroz)
```

<br>

```{r}
mroz1 = mroz[mroz$inlf == 1,]
```

<br>

```{r}
tab_model(lm(log(wage) ~ educ + exper + I(exper^2), data = mroz1))
```

<br>
<br>
<br>

```{r}
tab_model(lm(educ ~ motheduc + exper + I(exper^2), data = mroz1))
```

<br>
<br>
<br>

### Predict values of education by mother's education

```{r}
mod1 = lm(educ ~ motheduc + exper + I(exper^2), data = mroz1)
mroz1$educf = fitted(mod1)
```

<br>
<br>
<br>

### Fit a new model

```{r}
tab_model(lm(log(wage) ~ educf + exper + I(exper^2), data = mroz1))
```

<br>
<br>
<br>

```{r}
mod3 = ivreg(log(wage) ~ educ + exper + I(exper^2)|exper + I(exper^2) + motheduc, data = mroz1)
summary(mod3, diagnostics = T)
```

