---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<span class="h1">Logistic Regression for Multinomial and Ordinal responses</span>

<span class="h5">
Modern Social Analysis, HSE SPb<br>
*Course:* Quantitative Data Analysis
</span>

<span class="h4">
Vasily Khodakovsky</span>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

```


```{r include=FALSE}
#{r, results = "hide"}

Sys.setlocale("LC_ALL", "English")
```

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

# Data

```{r}
library(foreign)

data = read.dta("https://stats.idre.ucla.edu/stat/data/hsbdemo.dta")

options(width = 1000)
tail(data)
```

<br>
<br>
<br>
<br>
<br>

# Multinomial regression

```{r}
#treatment contrasts

library(sjPlot)
library(nnet)
m1 = multinom(prog ~ ses + math, data = data)
```

<br>

```{r}
summary(m1, Wald = T)
```


Socio-economic status doesn't affect the choice of a program.

1 point increse in math results increases the log of odds of choosing an academic program vs a general program by 0.082, and decreases the log of odds of choosing a vocation program vs a general program by 0.071.

1 point increse in math results increases odds of choosing academic program vs general probram by 9%, and decreases the log of odds of choosing vocation program vs general program by 7%.

<br>
<br>
<br>

```{r}
exp(coef(m1))
```

<br>
<br>
<br>

```{r}
library(stargazer)
stargazer(m1, type = "text")
```

<br>
<br>
<br>

```{r}
library(pscl)
pR2(m1) #bearable
```

<br>
<br>
<br>

```{r}
hitmiss(m1) #bearable
```

<br>
<br>
<br>

```{r}
plot_model(m1, type = "pred", terms = "math")
```

<br>
<br>
<br>

### Changing reference category

```{r}
data$prog2 = relevel(data$prog, ref = "academic")
m2 = multinom(prog2 ~ ses + math, data = data)
```

<br>

```{r}
summary(m2, Wald = T)
```

<br>
<br>
<br>

```{r}
stargazer(m2, type = "text")
```

<br>
<br>
<br>

### Dummy coding

```{r}
library(car)
data$voc = recode(data$prog, ' "vocation" = 1; else = 0 ')
table(data$prog)
```

<br>

```{r}
table(data$voc)
```

<br>

```{r}
data$gen = recode(data$prog, ' "general" = 1; else = 0 ')
table(data$prog)
```

<br>

```{r}
table(data$gen)
```

<br>

```{r}
data$ac = recode(data$prog, ' "academic" = 1; else = 0 ')
table(data$prog)
```

<br>

```{r}
table(data$ac)
```

<br>
<br>
<br>

### Binary regressions

```{r}
m_voc = glm(voc ~ ses + math, data = data, family = 'binomial')
m_gen = glm(gen ~ ses + math, data = data, family = 'binomial')
m_ac = glm(ac ~ ses + math, data = data, family = 'binomial')
```

<br>
<br>
<br>

```{r}
stargazer(m_voc, m_gen, m_ac, type = "text")
```

<br>

```{r}
exp(coef(m_voc))
```

<br>

```{r}
exp(coef(m_gen))
```

<br>

```{r}
exp(coef(m_ac))
```

<br>
<br>
<br>

```{r}
summary(m_voc)
```

<br>

```{r}
summary(m_gen)
```

<br>

```{r}
summary(m_ac)
```

<br>

Socio-economic status still doesn't affect the choice of a program.

1 point increase in math results decrease the log of odds of choosing a vocation program by 0.122 and increase the log of odds of choosing an academic program by 0.113.

1 point increase in math results decrease the odds of choosing a vocation program by 12% and increase the odds of choosing an academic program by 12%.

<br>
<br>
<br>

```{r}
pR2(m_voc)
```

<br>

```{r}
pR2(m_gen)
```

<br>

```{r}
pR2(m_ac)
```

<br>
<br>
<br>

```{r}
hitmiss(m_voc)
```

<br>

```{r}
hitmiss(m_gen)
```

<br>

```{r}
hitmiss(m_ac)
```

<br>
<br>
<br>

# Ordered Regression

```{r}
library(carData)
data(WVS)
```

<br>
<br>
<br>

```{r}
library(MASS)
m3 = polr(poverty ~ religion + degree + age + gender + country, data = WVS)
tab_model(m3)
```

<br>

The log of odds of more more positive evaluation of government efforts is 0.18 greater for the respondents who identify with a religious denomination.

The log of odds of more more positive evaluation of government efforts is 0.32 lower for the respondents who live in Norway compared to those who live in Australia.

With each year of age the log of odds of more more positive evaluation of government efforts increases by 0.01.

The odds of more positive evaluation of government efforts are 20% higher for the respondents who identify with a religious denomination.

The odds of more more positive evaluation of government efforts are 28% lower for the respondents who live in Norway compared to those who live in Australia.

With each year of age the odds of more more positive evaluation of government efforts increase by 1%.

<br>
<br>
<br>

```{r}
plot_model(m3, type = "pred", terms = "age")
```

<br>
<br>
<br>

```{r}
pR2(m3)
```


<br>
<br>
<br>

```{r}
hitmiss(m3)
```

