---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

<br>

<div style="text-align: center;">
<span class="h1">Poisson Models</span>

<span class="h5">
Modern Social Analysis, HSE SPb<br>
*Course:* Quantitative Data Analysis
</span>

<span class="h4">
Vasily Khodakovsky</span>
</div>

```{r include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

# Packages

```{r}
library(car)
library(sjPlot)
library(pscl)
library(texreg)
library(AER)
library(vcdExtra)
```

<br>
<br>
<br>
<br>
<br>


# Data

```{r}
head(bioChemists)
```

<br>
<br>
<br>
<br>
<br>

# Analysis

```{r}
mpois = glm(art ~ fem + mar + kid5 + phd + ment, data = bioChemists, family = poisson)
tab_model(mpois)
```

<br>
<br>
<br>

```{r}
mpois_inter = glm(art ~ mar + kid5 + phd*fem + ment, data = bioChemists, family = poisson)
tab_model(mpois_inter)
```

<br>
<br>
<br>
<br>
<br>

### Check for overdispersion

<br>

#### Checking linear relationship:

```{r}
dispersiontest(mpois, trafo = 1)
```

<br>

If p-value is < 0.05, them mean and variance are not equal.

Overdispersion is present - alpha (coef between mean and variance) is lower 1.

<br>
<br>
<br>

#### Checking quadratic relationship:

```{r}
dispersiontest(mpois, trafo = 2)
```

<br>

If p-value is < 0.05, it is quadratic.

Negative Binomial is more suitable.

<br>
<br>
<br>
<br>
<br>

### Quasi-Poisson

```{r}
fm_qpois <- glm(art ~ fem + mar + kid5 + phd + ment,
                data = bioChemists, family = quasipoisson)
```

<br>
<br>
<br>

```{r}
fm_nb <- MASS::glm.nb(art ~ fem + mar + kid5 + phd + ment, 
                      data = bioChemists)
```

<br>
<br>
<br>

```{r}
screenreg(list(mpois, fm_qpois, fm_nb))
```

<br>

Intercept and married people lost their significance.

By AIC, BIC, Log Likelihood - model3 is better.

Deviance of model1 and model2 are almost similar.

<br>
<br>
<br>
<br>
<br>

### Check for inflations of zeroes

```{r}
zero.test(table(bioChemists$art))
```

<br>

There are too many zeros; we need to use Zeroes Inflation Measures.

<br>
<br>
<br>

```{r}
fm_zinb1 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment|1,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb1)
```

<br>
<br>
<br>

```{r}
fm_zinb1 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb1)
```

<br>
<br>
<br>

```{r}
fm_zinb3 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment|ment,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb3)
```

<br>
<br>
<br>

```{r}
pR2(mpois)[4]
```

<br>

```{r}
pR2(fm_zinb3)[4]
```

<br>
<br>
<br>
<br>
<br>

```{r}
library(MASS)
data("Insurance")
head(Insurance)
```

<br>
<br>
<br>

```{r}
mod1 = glm(Claims ~ District + Group + Age, family = poisson, data = Insurance)

mod2 = glm(Claims ~ District + Group + Age + offset(log(Holders)),
           family = poisson, data = Insurance)

screenreg(list(mod1, mod2))
```
