---
output: 
  html_document:
    code_folding: show
    theme: readable
    df_print: paged
editor_options:
    chink_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

<br>

<div style="text-align: center;">
<b><span class="h1">Poisson Models</span></b>

<span class="h5">
<b>Modern Social Analysis, HSE SPb</b><br>
*Course:* <b>Quantitative Data Analysis</b>
</span>

<span class="h4"><b>
Vasily Khodakovsky</span>
</b></div>

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

# Packages

```{r}
library(car)
library(sjPlot)
library(pscl)
library(texreg)
library(AER)
library(vcdExtra)
```

<br>
<br>
<br>
<br>
<br>


# Data

```{r}
head(bioChemists)
```

<br>
<br>
<br>
<br>
<br>

# Analysis

```{r}
mpois = glm(art ~ fem + mar + kid5 + phd + ment, data = bioChemists, family = poisson)
tab_model(mpois)
```

<br>
<br>
<br>

```{r}
mpois_inter = glm(art ~ mar + kid5 + phd*fem + ment, data = bioChemists, family = poisson)
tab_model(mpois_inter)
```

<br>
<br>
<br>
<br>
<br>

### Check for overdispersion

```{r}
dispersiontest(mpois, trafo = 1) #checks linear relationship
```

P-value < 0.05 => mean and variance are not equal;

Overdispersion is present - alpha is lower 1 (coef between mean and variance).

<br>
<br>
<br>

```{r}
dispersiontest(mpois, trafo = 2) #checks quadratic relationship
```


If p-value < 0.05, it is quadratic.

Negative Binomial is more suitable.


<br>
<br>
<br>
<br>
<br>

### Quasi-Poisson

```{r}
fm_qpois <- glm(art ~ fem + mar + kid5 + phd + ment,
                data = bioChemists, family = quasipoisson)
```

<br>
<br>
<br>

```{r}
fm_nb <- MASS::glm.nb(art ~ fem + mar + kid5 + phd + ment, 
                      data = bioChemists)
```

<br>
<br>
<br>

```{r}
screenreg(list(mpois, fm_qpois, fm_nb))
```

Intercept and married people lost their significance.

By AIC, BIC, Log Likelihood - model3 is better.

Deviance of model1 and model2 are almost similar.

<br>
<br>
<br>
<br>
<br>

### Check for inflations of zeroes

```{r}
zero.test(table(bioChemists$art))
```

There are too many zeros; need to use Zeroes Inflation Measures.

There are two groups of people: who have and who don’t have articles.

<br>
<br>
<br>

```{r}
#by default it is poisson
#"|1" -> separation between count and zero models = > specify for whose who have articles
#we predict here zero, whether or not the value will be zero

#first part is negative binomial model, and the second part after "| 1"
#are those who will have zero published articles only for intercept

fm_zinb1 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment|1,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb1)
```

<br>

log(theta) is a coeff between variance and mean.

<br>
<br>
<br>

```{r}
fm_zinb1 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb1)
```

<br>

The log of odds do not have publications decreases by 0.88 with each mentor’s publication.

<br>
<br>
<br>

```{r}
fm_zinb3 <- zeroinfl(art ~ fem + mar + kid5 + phd + ment|ment,
                     data = bioChemists, dist = "negbin")

summary(fm_zinb3)
```

Prediction of success -> not publushing.

<br>
<br>
<br>

```{r}
pR2(mpois)[4]
```

<br>

```{r}
pR2(fm_zinb3)[4]
```

<br>
<br>
<br>
<br>
<br>

```{r}
library(MASS)
data("Insurance")
head(Insurance)
```

<br>
<br>
<br>

```{r}
#certain variable as exposure (GDP per capita, number a cigarettes by day/week)

mod1 = glm(Claims ~ District + Group + Age, family = poisson, data = Insurance)

mod2 = glm(Claims ~ District + Group + Age + offset(log(Holders)),
           family = poisson, data = Insurance)

screenreg(list(mod1, mod2))
```











