---
output: 
  html_document:
    code_folding: show
    theme: readable
    toc: true
    toc_float: true
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<span class="h1">Bipartite networks</span>

<span class="h5">
Modern Social Analysis, HSE SPb<br>
*Course:* Social Network Analysis</span>

<span class="h4">Vasily Khodakovsky</span>
</div>

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Commitees network

```{r}
df = read.csv("C:/Users/Vasily/Downloads/bimodal.csv", header = T, sep = ";")
mat = as.matrix(table(df))
```

<br>

#### Converting matrix to the bipartite network:

```{r}
library(igraph)
bg_com = graph.incidence(mat)
```

<br>

## Visualization

```{r}
shape = ifelse(V(bg_com)$type, "square", "circle") # assign shape by node type
col = ifelse(V(bg_com)$type, "tomato3", "lightblue") # assign color by node type

degree_com = degree(bg_com)
```

<br>

### Default layout

```{r, fig.height = 10, fig.width = 10}
plot(bg_com, 
     vertex.shape = shape, 
     vertex.color = col, 
     vertex.size = ifelse(shape=="square", degree_com/2, 4),
     vertex.label = ifelse(shape=="square", V(bg_com)$name, NA),
     vertex.label.color = "black",
     vertex.label.dist = 0.15,
     main = "Members of commitees network:")
```

<br>
<br>
<br>

### Bipartite layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(bg_com, layout = layout_as_bipartite(bg_com),
     vertex.shape = shape, 
     vertex.color = col, 
     vertex.size = ifelse(shape=="square", degree_com/2, 4),
     vertex.label = NA,
     main = "Members of commitees network\n(Bipartite layout):")
```

<br>
<br>
<br>

### Kamada-Kawai layout

<br>

```{r, fig.height = 10, fig.width = 10} 
plot(bg_com, layout = layout_with_kk(bg_com),
     vertex.shape = shape, 
     vertex.color = col, 
     vertex.size = ifelse(shape=="square", degree_com/2, 4),
     vertex.label = ifelse(shape=="square", V(bg_com)$name, NA),
     vertex.label.color = "black",
     vertex.label.dist = 0.15,
     main = "Members of commitees network\n(Kamada-Kawai layout):")
```

<br>
<br>
<br>
<br>

## Dividing the network

```{r}
m_row <- mat %*% t(mat) 
m_column <- t(mat) %*% mat
```

<br>

```{r} 
bg_row <- graph_from_adjacency_matrix(m_row,
                                      mode = "undirected",
                                      diag = FALSE,
                                      weighted = TRUE)
e <- E(bg_row)$weight
```

<br>

```{r} 
bg_column <- graph_from_adjacency_matrix(m_column,
                                         mode = "undirected",
                                         diag = FALSE, 
                                         weighted = TRUE)
e1 <- E(bg_column)$weight
```

<br>
<br>
<br>

### Network of committee members

<br>

```{r, fig.height = 10, fig.width = 10} 
plot(bg_row, layout = layout_with_fr(bg_row),
     vertex.shape = shape,
     vertex.label = NA,
     vertex.color = col, vertex.size = 5,
     edge.width = e)
```

<br>
<br>
<br>

### Network of committees

<br>

```{r, fig.height = 10, fig.width = 10} 

plot(bg_column, 
     vertex.shape = "square",
     vertex.label.color = "black",
     vertex.color = "tomato3", vertex.size = 27.5,
     edge.width = e1)
```

<br>
<br>
<br>
<br>
<br>

# Scrapped data

<br>

### Data collection

```{r}
library(rvest)
```

<br>

```{r include=FALSE}
url <- "https://ru.wikipedia.org/wiki/%D0%90%D0%B4%D0%BC%D0%B8%D0%BD%D0%B8%D1%81%D1%82%D1%80%D0%B0%D1%82%D0%B8%D0%B2%D0%BD%D0%BE-%D1%82%D0%B5%D1%80%D1%80%D0%B8%D1%82%D0%BE%D1%80%D0%B8%D0%B0%D0%BB%D1%8C%D0%BD%D0%BE%D0%B5_%D0%B4%D0%B5%D0%BB%D0%B5%D0%BD%D0%B8%D0%B5_%D0%A1%D0%B0%D0%BD%D0%BA%D1%82-%D0%9F%D0%B5%D1%82%D0%B5%D1%80%D0%B1%D1%83%D1%80%D0%B3%D0%B0#%D0%92%D0%BD%D1%83%D1%82%D1%80%D0%B8%D0%B3%D0%BE%D1%80%D0%BE%D0%B4%D1%81%D0%BA%D0%B8%D0%B5_%D0%BC%D1%83%D0%BD%D0%B8%D1%86%D0%B8%D0%BF%D0%B0%D0%BB%D1%8C%D0%BD%D1%8B%D0%B5_%D0%BE%D0%B1%D1%80%D0%B0%D0%B7%D0%BE%D0%B2%D0%B0%D0%BD%D0%B8%D1%8F_%D0%A1%D0%B0%D0%BD%D0%BA%D1%82-%D0%9F%D0%B5%D1%82%D0%B5%D1%80%D0%B1%D1%83%D1%80%D0%B3%D0%B0"
```

#### Used website: Wikipedia page about administrative division of Saint-Petersburg

```{r}
page <- read_html(url)
```

<br>

#### Extracting municipalities' names:

```{r}
municipality <- page %>%
  html_nodes("table.standard tbody tr td:nth-child(5) a") %>%
  html_text()
```

<br>

#### Extracting districts' names:

```{r}
district <- page %>%
  html_nodes("table.standard tbody tr td:nth-child(10) a") %>%
  html_text()
```

<br>
<br>
<br>

### Network creation

<br> 

#### Transormation data to the network:

```{r}
net <- data.frame(Column1 = district, Column2 = municipality)
tnet <- table(net)
matrix <- as.matrix(tnet)
gr <- graph.incidence(matrix)
```

<br>
<br>
<br>

### Administrative division of Saint-Petersburg

<br>

```{r}
col = ifelse(V(gr)$type, "wheat", "lightblue")
shape = ifelse(V(gr)$type, "square", "circle")
size = degree(gr, mode = "in")
```

<br>

```{r, fig.height = 10, fig.width = 10} 
plot(gr, layout = layout.fruchterman.reingold(gr),
     edge.arrow.size = 1,
     vertex.size = size*1.4,
     vertex.shape = shape,  
     vertex.color = col,
     vertex.frame.color = col,
     vertex.label.cex = ifelse(shape=="circle", 0.6, 0.3),
     vertex.label.color = "black")
```


<br>
<br>
<br>

```{r, fig.height = 10, fig.width = 10}
cl_louv <- cluster_louvain(gr)
par(mar = c(0,0,0,0))
plot(cl_louv, 
     gr,
     mark.groups = communities(cl_louv),
     col = membership(cl_louv),
     edge.color = "white",
     vertex.size = 2,
     vertex.shape = shape,
     vertex.frame.color = membership(cl_louv),
     vertex.label = ifelse(shape=="circle", V(gr)$name, NA),
     vertex.label.color = 'black',
     vertex.label.cex = 0.8)
```

