---
output: 
  html_document:
    code_folding: show
    theme: readable
    toc: true
    toc_float: true
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div style="text-align: center;">
<span class="h1"><b>Class networks</b></span>

<span class="h5">
<b>Modern Social Analysis, HSE SPb</b><br>
*Course:* <b>Social Network Analysis</b></span>

<span class="h4"><b>Vasily Khodakovsky</b></span>
</div>

<br>
<br>
<br>
<br>
<br>

<br>
<br>
<br>
<br>
<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# General information

```{r}
library(igraph)
help <- read_graph("C:/Users/Vasily/Downloads/help_asking.net", format = "pajek")
friends <- read_graph("C:/Users/Vasily/Downloads/positive.net", format = "pajek")
negative <- read_graph("C:/Users/Vasily/Downloads/negative.net", format = "pajek")
```

<br>
<br>
<br>

### Nodes

```{r}
library(kableExtra)
V(help) %>% 
  as.data.frame() %>% 
  setNames(c("Ordinal number")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

### Edges

```{r}
E(help)
```

<br>
<br>
<br>

### Density
 
 Density captures how many edges there are in a network divided by the total possible number of edges.

```{r}
data.frame(graph.density(help), graph.density(friends), graph.density(negative)) %>% 
  as.data.frame() %>% 
  setNames(c("Density of `help`", "Density of `friends`", "Density of `negative`")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```
 
<br>
<br>
<br>
 
### Degree centrality
 
 Degree centrality of a node shows the number of edges of its node. 
 
 In-degree is the number of in-coming links (predecessor nodes). 
 
 Out-degree is the number of out-going links (successor nodes).

```{r}
help_degree = degree(help, v = V(help), mode = "in")
friends_degree_in = degree(friends, v = V(friends), mode = "in")
friends_degree_out = degree(friends, v = V(friends), mode = "out")
negative_degree = degree(negative, v = V(negative), mode = "in")
```

<br>

```{r}
data.frame(help = help_degree, friends_in = friends_degree_in, 
           friends_out = friends_degree_out, negative = negative_degree) %>% 
  as.data.frame() %>% 
  setNames(c("`help` degree", "`friends` in-degree", 
             "`friends` out-degree", "`negative` degree")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
data.frame(mean(help_degree), mean(friends_degree_in),
           mean(friends_degree_out), mean(negative_degree)) %>% 
  as.data.frame() %>% 
  setNames(c("Mean degree of `help`", "Mean in-degree of `friends`", 
             "Mean out-degree of `friends`", "Mean degree of `negative`")) %>% 
  round(., digits = 2) %>%
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```
 
<br>
<br>
<br>
<br>
<br>
 
# Visualization

<br>

### Help network
 
```{r, fig.height = 10, fig.width = 10}
plot(help, 
     edge.arrow.size = 0.3, 
     vertex.label = V(help)$name, 
     vertex.size = help_degree*3)
```
 
<br>
<br>
<br>

### Friends network

<br>
 
```{r, fig.height = 10, fig.width = 10}
plot(friends, 
     edge.arrow.size = 0.3, 
     vertex.label = V(friends)$name, 
     vertex.size = friends_degree_in*3)
```
 
<br>
<br>
<br>

### Negative network

<br>
 
```{r, fig.height = 10, fig.width = 10}
plot(negative, 
     edge.arrow.size = 0.3, 
     vertex.label = V(negative)$name, 
     vertex.size = negative_degree*3)
```
 
<br>
<br>
<br>
<br>
<br>
 
### `neighbors` function
 
```{r}
neigh.savva <- neighbors(negative, #name of the network
                         V(negative)[name=="Savva"], #name of the node
                         mode = "in") #mode: in (incoming), out (outcoming), all (both)

V(negative)$color <- "white"
V(negative)[neigh.savva]$color <- "orange"
V(negative)["Savva"]$color <- "red"

plot(negative,
     edge.arrow.size = .2,
     vertex.size = 30,
     vertex.label.cex = .6)
```

<br>

Savva's node is red; nodes of classmates who doesn't like him are orange. 

<br>
<br>
<br>

```{r}
neigh.Lisa <- neighbors(help,
                        V(help)[name=="Lisa"],
                        mode = "in")

V(help)$color <- "white"
V(help)[neigh.Lisa]$color <- "skyblue"
V(help)["Lisa"]$color <- "palegreen"

plot(help,
     edge.arrow.size = .2,
     vertex.size = 30,
     vertex.label.cex = .6)
```

<br>

Lisa's node is blue; node of each person who asks her for help is green.

<br>
<br>
<br>
<br>
<br>

# Community detection

```{r}
friendsUn <- as.undirected(friends, mode = "mutual")
```

<br>

To perform some algorithms, it is necessary to transform network to undirected.

<br>
<br>
<br>

### Cluster Louvian method

```{r}
cl_louv <- cluster_louvain(friendsUn)
```

<br>

```{r}
modularity(cl_louv) %>% 
  setNames(c("Modulatity (the strength of the community formation)")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
sizes_cl <- data.frame(sizes(cl_louv))
sizes_cl %>% 
  as.data.frame() %>% 
  setNames(c("Ordinal number of the community", "Number of members")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")

```

<br>

```{r, fig.height = 10, fig.width = 10}
plot(cl_louv,
     friendsUn,
     col = membership(cl_louv),
     mark.groups = communities(cl_louv),
     edge.color = "grey",
     vertex.label = V(friendsUn)$name,
     vertex.label.cex = .6)
```

<br>
<br>
<br>

### Edge betweenness method


```{r}
edge_between <- edge.betweenness.community(friendsUn)
```

<br>

```{r}
modularity(edge_between) %>% 
  setNames(c("Modulatity")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

```{r}
sizes_eb <- data.frame(sizes(edge_between))
sizes_eb %>% 
  as.data.frame() %>% 
  setNames(c("Ordinal number of the community", "Number of members")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

```{r, fig.height = 10, fig.width = 10}
plot(edge_between,
     friendsUn,
     col = membership(edge_between),
     mark.groups = communities(edge_between),
     edge.color = "grey",
     vertex.label = V(friendsUn)$name,
     vertex.label.cex = .6)
```

<br>
<br>
<br>

### Fast Greedy algorithm


```{r}
fast_greedy <- fastgreedy.community(friendsUn)
```

<br>

```{r}
modularity(fast_greedy) %>% 
  setNames(c("Modulatity")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

```{r}
sizes_fg <- data.frame(sizes(edge_between))
sizes_fg %>% 
  as.data.frame() %>% 
  setNames(c("Ordinal number of the community", "Number of members")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")

```

<br>
<br>
<br>


```{r, fig.height = 10, fig.width = 10}
plot(fast_greedy,
     friendsUn,
     col = membership(fast_greedy),
     mark.groups = communities(fast_greedy),
     edge.color = "grey",
     vertex.label = V(friendsUn)$name,
     vertex.label.cex = .6)
```


<br>
<br>
<br>
<br>
<br>

# Layouts

<br>

#### Types of layout:

```{r}
r <- layout_randomly(friends, dim = 2) #only random layout needs "dim" 
c <- layout_in_circle(friends)
gr <- layout_on_grid(friends)
kk <- layout.kamada.kawai(friends)
fr <- layout_with_fr(friends)
```

<br>
<br>
<br>

#### Networks' settings:

```{r}
my_resolution = 100
my_palette = colorRampPalette(c('blue','lightgreen'))

degree = degree(friends, v = V(friends), mode = "in")
my_colors = my_palette(my_resolution)[as.numeric(cut(degree, breaks = my_resolution))]
```

<br>
<br>
<br>

### Random layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(friends, layout = r, 
     vertex.label = V(friends)$name, 
     edge.arrow.size = 0.3, 
     vertex.size = degree*4,
     vertex.label.cex = 0.6,
     vertex.label.color = 'black',
     vertex.color = my_colors)
```

<br>
<br>
<br>

### Circle layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(friends, layout = c, 
     vertex.label = V(friends)$name, 
     edge.arrow.size = 0.3, 
     vertex.size = degree*4,
     vertex.label.cex = 0.6,
     vertex.label.color = 'black',
     vertex.color = my_colors)
```

<br>
<br>
<br>

### Grid layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(friends, layout = gr, 
     vertex.label = V(friends)$name, 
     edge.arrow.size = 0.3, 
     vertex.size = degree*4,
     vertex.label.cex = 0.6,
     vertex.label.color = 'black',
     vertex.color = my_colors)
```

<br>
<br>
<br>

### Kamada-Kawai layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(friends, layout = kk, 
     vertex.label = V(friends)$name, 
     edge.arrow.size = 0.3, 
     vertex.size = degree*4,
     vertex.label.cex = 0.6,
     vertex.label.color = 'black',
     vertex.color = my_colors)
```

<br>
<br>
<br>

### Fruchteman-Reingold layout

<br>

```{r, fig.height = 10, fig.width = 10}
plot(friends, layout = fr, 
     vertex.label = V(friends)$name, 
     edge.arrow.size = 0.3, 
     vertex.size = degree*4,
     vertex.label.cex = 0.6,
     vertex.label.color = 'black',
     vertex.color = my_colors)
```
