---
output: 
  html_document:
    code_folding: show
    theme: readable
    toc: true
    toc_float: true
    df_print: paged
editor_options:
    chink_output_type: console
---

<br>

<div class="h1" style="text-align: center;">
Relationship between Europeans' trust in political institutions and their attitude to government surveillance
</div>

<div class="h5" style="text-align: center;">
*Modern Social Analysis, HSE SPb (Course Thesis)*
</div>

<div class="h4" style="text-align: center;">
Vasily Khodakovsky
</div>

<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```
 
  
```{r, include = FALSE}
Sys.setlocale("LC_ALL", "English")
```
  
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

# Data

<br>

#### [5th wave of European Values Study (EVS)](https://europeanvaluesstudy.eu/)

<br>

```{r packages}
library(plyr)
library(dplyr)
library(kableExtra)
```
 
<br>

```{r data frame}
library(foreign)
EVS_raw <- read.spss("C:/Users/Vasily/Downloads/EVS.sav", 
                     use.value.labels = T, to.data.frame = T)
```
 
<br>
<br>
<br>
<br>
<br>

### Countries


```{r}
table(EVS_raw$country) %>% 
  as.data.frame() %>% 
  setNames(c("Countries", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```


<br>
<br>
<br>

#### Creating new variable `part`, that indicates is country ex-socialist or not (Eastern Europe or not):

```{r}
EVS_raw$part <- "West"

countries_part <- c("Albania", "Armenia", "Azerbaijan", 
                    "Belarus", "Bosnia and Herzegovina", "Bulgaria", 
                    "Croatia", "Czechia", "Estonia", 
                    "Georgia", "Hungary", "Lithuania", 
                    "Montenegro", "North Macedonia", "Poland", 
                    "Romania", "Russia", "Serbia", 
                    "Slovakia", "Slovenia")

EVS_raw <- EVS_raw %>% 
  mutate(part = ifelse(country %in% countries_part, "East", part))


EVS_raw$part = as.factor(EVS_raw$part)
```

<br>

```{r}
table(EVS_raw$part) %>% 
  as.data.frame() %>% 
  setNames(c("Part of Europe", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```


<br>
<br>
<br>

### Trust in political institutions 

<br>

#### Q38G) Parliament:

```{r}
table(EVS_raw$v121) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in parliament", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v121 <- dplyr::recode(EVS_raw$v121, 
                              "a great deal" = 4,
                              "quite a lot" = 3,
                              "not very much" = 2,
                              "none at all" = 1)
```

<br>

```{r}
EVS_raw$v121 <- as.numeric(EVS_raw$v121)
```

<br>

```{r}
table(EVS_raw$v121) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in parliament", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Q38H) Civil service:

```{r}
table(EVS_raw$v122) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in civil service", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v122 <- dplyr::recode(EVS_raw$v122, 
                              "a great deal" = 4,
                              "quite a lot" = 3,
                              "not very much" = 2,
                              "none at all" = 1)
```

<br>

```{r}
EVS_raw$v122 <- as.numeric(EVS_raw$v122)
```

<br>

```{r}
table(EVS_raw$v122) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in civil service", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```


<br>
<br>
<br>

#### Q38Q) Government:

```{r}
table(EVS_raw$v131) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in government", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v131 <- dplyr::recode(EVS_raw$v131, 
                              "a great deal" = 4,
                              "quite a lot" = 3,
                              "not very much" = 2,
                              "none at all" = 1)
```

<br>

```{r}
EVS_raw$v131 <- as.numeric(EVS_raw$v131)
```

<br>

```{r}
table(EVS_raw$v131) %>% 
  as.data.frame() %>% 
  setNames(c("Level of confidence in government", "Number of respondents")) %>% 
  kbl() %>% 
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Index:

Creating Political Trust Index

```{r}
EVS_raw$polit <- (EVS_raw$v121 + EVS_raw$v122 + EVS_raw$v131)/3
```

<br>
<br>
<br>

##### Cronbach's alpha of the index

```{r}
library(ltm)
polit_cronbach <- EVS_raw %>% dplyr::select(v121, v122, v131)
cronbach.alpha(polit_cronbach, na.rm = T)
```

Level of Cronbach's alpha is fine.

<br>
<br>
<br>
<br>
<br>


### Surveillance

<br>

#### Q58A) Should the government have the right to keep people under video surveillance in public areas:

```{r}
table(EVS_raw$v205) %>%
  as.data.frame() %>%
  setNames(c("Attitude to the video surveillance", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v205 <- dplyr::recode(EVS_raw$v205, 
                              "definitely should have the right" = 4,
                              "probably should have the right" = 3,
                              "probably should not have the right" = 2,
                              "definitely should not have the right" = 1)
```

<br>

```{r}
EVS_raw$v205 <- as.numeric(EVS_raw$v205)
```

<br>

```{r}
table(EVS_raw$v205) %>%
  as.data.frame() %>%
  setNames(c("Attitude to the video surveillance", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Q58B) Should the government have the right to monitor all e-mails and any other information exchanged on the Internet:

```{r}
table(EVS_raw$v206) %>%
  as.data.frame() %>%
  setNames(c("Attitude to the internet surveillance", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v206 <- dplyr::recode(EVS_raw$v206, 
                              "definitely should have the right" = 4,
                              "probably should have the right" = 3,
                              "probably should not have the right" = 2,
                              "definitely should not have the right" = 1)
```

<br>

```{r}
EVS_raw$v206 <- as.numeric(EVS_raw$v206)
```

<br>

```{r}
table(EVS_raw$v206) %>%
  as.data.frame() %>%
  setNames(c("Attitude to the internet surveillance", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Q58C) Should the government have the right to collect information about anyone living in the country without their knowledge:

```{r}
table(EVS_raw$v207) %>%
  as.data.frame() %>%
  setNames(c("Attitude to collecting information about anyone", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

```{r}
EVS_raw$v207 <- dplyr::recode(EVS_raw$v207, 
                              "definitely should have the right" = 4,
                              "probably should have the right" = 3,
                              "probably should not have the right" = 2,
                              "definitely should not have the right" = 1)
```

<br>

```{r}
EVS_raw$v207 <- as.numeric(EVS_raw$v207)
```

<br>

```{r}
table(EVS_raw$v207) %>%
  as.data.frame() %>%
  setNames(c("Attitude to the collecting information about anyone", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Index:

Creating Surveillance Attitude Index

```{r}
EVS_raw$surv <- (EVS_raw$v205 + EVS_raw$v206 + EVS_raw$v207)/3
```

<br>

##### Cronbach's alpha of the index

```{r}
surv_cronbach <- EVS_raw %>% dplyr::select(v205, v206, v207)
cronbach.alpha(surv_cronbach, na.rm = T)
```

Level of Cronbach's alpha is fine.

<br>
<br>
<br>
<br>


### Gender


```{r list of variablesw2}
table(EVS_raw$v225) %>%
  as.data.frame() %>%
  setNames(c("Gender", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

#### Renaming variable:

```{r}
EVS_raw$gender <- EVS_raw$v225
```

<br>
<br>
<br>
<br>
<br>

### Education

```{r list of variablesvcc}
table(EVS_raw$v243_r) %>%
  as.data.frame() %>%
  setNames(c("Education category", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

#### Deleting "other" education because it has small proportion:

```{r}
EVS_raw <- EVS_raw %>% filter(v243_r == "lower" |
                              v243_r == "medium" | 
                              v243_r == "higher")
EVS_raw <- EVS_raw[!EVS_raw$v243_r == "other",]
EVS_raw$v243_r = droplevels(EVS_raw$v243_r) 
```

<br>

```{r}
table(EVS_raw$v243_r) %>%
  as.data.frame() %>%
  setNames(c("Education category", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

#### Renaming variable:

```{r}
EVS_raw$education <- EVS_raw$v243_r
```

<br>
<br>
<br>
<br>
<br>

### Income


Three categories, depend on the income in the country

```{r list of variablesdls}
table(EVS_raw$v261_r) %>%
  as.data.frame() %>%
  setNames(c("Income category", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

#### Renaming variable:

```{r}
EVS_raw$income <- EVS_raw$v261_r
```

<br>
<br>
<br>
<br>
<br>

### Size of town


```{r list sdfof variables}
table(EVS_raw$v276_r) %>%
  as.data.frame() %>%
  setNames(c("Size of town", "Number of respondents")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

#### Renaming variable:

```{r}
EVS_raw$size_town <- EVS_raw$v276_r
```

<br>
<br>
<br>
<br>
<br>

### Age

```{r listfg of variables}
data.frame(c("Min", "Max"),
           c(head(levels(EVS_raw$age))[1], tail(levels(EVS_raw$age))[6])) %>%
  as.data.frame() %>%
  setNames(c("Age", " ")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

```{r age}
EVS_raw$age = mapvalues(EVS_raw$age, from = c("82 and older"), to = c("82")) 
detach("package:plyr", unload=TRUE)

EVS_raw$age <-as.numeric(as.character(EVS_raw$age))

summary(EVS_raw$age)
```

<br>
<br>
<br>
<br>
<br>


### Final dataset


```{r check}
EVS <- EVS_raw %>% dplyr::select(country, part, 
                                 polit, surv, 
                                 gender, age, 
                                 education, income, 
                                 size_town)

summary(EVS)
```


<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

  
  
# Exploratory Analysis

<br>

### Descriptive Statistics

```{r, fig.height = 10, fig.width = 10}
par(mfrow = c(2,2))
plot(EVS$gender, col = "lightblue1", main = "Gender")
plot(EVS$income, col = "lightblue2", main = "Income")
plot(EVS$education, col = "lightblue3", main = "Education")
plot(EVS$size_town, col = "lightblue4", 
     names.arg = c("<5k", "5k-20k", "20k-100k", "100k-500k", "500k+"), 
     las = 2, 
     cex.names = 0.8,
     main = "Size of town",)
```

<br>
<br>
<br>

```{r}
library(psych)
round(describe(EVS[c("polit", "surv", "age")]),2) %>% 
  dplyr::select(min, max, mean, median, sd) %>%
  as.data.frame() %>%
  setNames(c("Min", "Max", "Mean", "Median", "SD")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>
<br>
<br>

#### Age:

```{r, fig.width = 10}
par(mfrow = c(1,2))
hist(EVS$age, col = "tomato3", xlab = "Age", main = "Histogram of age")
hist(log(EVS$age), col = "tomato4", xlab = "log(age)", main = "Histogram of log(age)")
```

<br>

```{r}
summary(EVS$age)
```

<br>

```{r}
library(car)
qqPlot(EVS$age, xlab = "Norm quantiles", ylab = "Age", main = "QQ-Plot of age")
```

<br>
<br>
<br>

#### Surveillance Attitude Index and Political Trust Index:

```{r}
par(mfrow = c(1, 2))

boxplot(EVS$surv, col = "wheat1", main = "Boxplot of the\nSurveillance Attitude Index")
mean_surv <- mean(EVS$surv, na.rm = T)
text(x = 1, y = mean_surv, labels = paste0(round(mean_surv, 2)), pos = 3)

boxplot(EVS$polit, col = "wheat2", main = "Boxplot of the\nPolitical Trust Index")
mean_polit <- mean(EVS$polit, na.rm = T)
text(x = 1, y = mean_polit, labels = paste0(round(mean_polit, 2)), pos = 3)
```

<br>
<br>
<br>
<br>
<br>

### Surveillance Attitude map

```{r}
averages <- aggregate(surv ~ country, data = EVS, FUN = function(x) round(mean(x), 2)) 
averages %>% 
  setNames(c("Country", "Surveillance Attitude Index")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")
```

<br>

```{r, fig.height = 10, fig.width = 10}
library(rnaturalearth)
world <- ne_countries(scale = "medium", returnclass = "sf")
colnames(averages)[1] <- "sovereignt"
averages$sovereignt <- dplyr::recode(averages$sovereignt, 
                              "Czechia" = "Czech Republic",
                              "North Macedonia" = "Macedonia",
                              "Great Britain" = "United Kingdom")

world <- merge(world, averages, by = "sovereignt")
```

<br>

```{r, fig.height = 10, fig.width = 10}
library(ggplot2)
ggplot(data = world) +
  geom_sf(aes(fill = surv)) +
  coord_sf(xlim = c(-40, 51), ylim = c(35.7, 72), expand = FALSE) +
  scale_fill_gradient(low = "tomato2", high = "lightblue2") +
  labs(fill = "Surv") + ggtitle("Surveillance Attitude Index")
```

<br>
<br>
<br>
<br>
<br>

### Correlational Analysis

<br>

#### Scatter plots for dependent and independent variables:

```{r, fig.width = 10}
sca1 = ggplot(EVS, aes(x = age, y = surv)) + geom_point() + 
  xlab("Age") + ylab("Surveillance Attitude Index") +  
  geom_smooth() + theme_bw() 

sca2 = ggplot(EVS, aes(x = polit, y = surv)) + geom_point() + 
  xlab("Political Trust Index") + ylab("Surveillance Attitude Index") +  
  geom_smooth() + theme_bw() 

grid_sc <- cowplot::plot_grid(sca1, sca2)
cowplot::plot_grid(grid_sc, ncol = 1, rel_heights = c(0.1, 1)) 
```

<br>
<br>
<br>

#### Correlation:

```{r}
corr_data <- data.frame(EVS$surv, EVS$age, EVS$polit)
colnames(corr_data) <- c("Surveillance Attitude Index", "Age", "Political Trust Index")
cor <- cor(corr_data, method = "kendall", use = "pairwise.complete.obs")
```

<br>

```{r}
library(ggcorrplot)
ggcorrplot(cor, type = "lower", lab = TRUE)
```

<br>
<br>
<br>
<br>
<br>

### Statistical test for means' comparison

<br>

#### By Gender:
```{r} 
wilcox.test(EVS$surv ~ EVS$gender)
```

<br>

```{r} 
boxplot(EVS$surv~EVS$gender,
        col = "aquamarine2",
        xlab = "Gender", 
        ylab = "Index of Surveillance",
        main = "Comparison of mean values of the\n Surveillance Attitude Index by gender")

means <- aggregate(surv ~ gender, data = EVS, FUN = mean)
text(x = seq_along(means$gender), y = means$surv, labels = round(means$surv, 2),
     pos = 3, offset = 0.5)
```

<br>
<br>
<br>

#### By Income:

```{r}
kruskal.test(surv ~ income, data = EVS)
```

<br>

```{r}
EVS[c("surv", "income")] %>% na.omit() %>% 
ggplot() +
  geom_boxplot(aes(x = surv, y = income)) +
  xlab("Surveillance Attitude Index") + ylab("Income") + theme_bw()
```

<br>

```{r}
library(FSA)
dunnTest(EVS$surv,EVS$income, method = "bonferroni")
```

There is a statistical difference in mean values of Surveillance Attitude Index<br>
between groups with low and high income.

<br>

```{r}
EVS$income <- as.factor(EVS$income)
EVS %>% dplyr::group_by(income) %>% summarize(mean = mean(surv, na.rm = T)) %>%
  as.data.frame() %>%
  setNames(c("Income", "Mean of the Surveillance Attitude Index")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria") 
```

<br>
<br>
<br>

#### By Size of town:

```{r}
kruskal.test(surv ~ size_town, data = EVS)
```

<br>

```{r}
EVS[c("surv", "size_town")] %>% na.omit() %>% 
ggplot() +
  geom_boxplot(aes(x = surv, y = size_town)) +
  xlab("Surveillance Attitude Index") + ylab("Size of town") + theme_bw()
```

<br>

```{r} 
dunnTest(EVS$surv, EVS$size_town, method = "bonferroni")
```

<br>

```{r} 
EVS %>% dplyr::group_by(size_town) %>% summarize(mean = mean(surv, na.rm = T)) %>%
  as.data.frame() %>%
  setNames(c("Size of town", "Mean of the Surveillance Attitude Index")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")  
```

<br>
<br>
<br>

#### By Education:

```{r}
kruskal.test(surv ~ education, data = EVS)
```

<br>

```{r}
EVS[c("surv", "education")] %>% na.omit() %>% 
ggplot() +
  geom_boxplot(aes(x = surv, y = education)) +
  xlab("Surveillance Attitude Index") + ylab("Education") + theme_bw()
```

<br>

```{r} 
dunnTest(EVS$surv, EVS$education, method = "bonferroni")
```

<br>

```{r} 
EVS %>% dplyr::group_by(education) %>% summarize(mean = mean(surv, na.rm = T)) %>%
  as.data.frame() %>%
  setNames(c("Education", "Mean of the Surveillance Attitude Index")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria")   
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

# Hypotheses testing

<br>

#### H1: The level of political trust is positively associated with the approval of government surveillance;

#### H2: The level of political trust has a stronger influence on the approval of surveillance in post-communist countries than in Western European countries;

#### H3: Age has a stronger influence on the approval of surveillance in Western European countries than in post-communist countries.

<br>
<br>
<br>

### Models 1 & 2

```{r}
library(sjPlot)
m1 <- lm(surv ~ polit, data = EVS)
tab_model(m1)
```

<br>

```{r}
m2 <- lm(surv ~ polit*part, data = EVS)
tab_model(m2)
```

<br>

```{r}
library(interactions)
interact_plot(m2, pred = polit, modx = part, plot.points = F) 
```

<br>
<br>
<br>

#### Check for multicollinearity:

```{r} 
vif(m2) %>%
  as.data.frame() %>%
  setNames(c("VIF")) %>%
  kbl() %>%
  kable_classic(html_font = "Cambria") 
```

There is a multicollinearity presence for interaction effect.

<br>
<br>
<br>

#### Leverage-Residual Plots:

```{r} 
plot(hatvalues(m1), rstudent(m1), type = 'n')
#thresholds for residuals

abline(h = c(-2, 2), lty = 2)
#thresholds for leverages

abline(v = c(2,3)*1/56407, lty = 2)

cook <- sqrt(cooks.distance(m1))
points(hatvalues(m1), rstudent(m1), cex = 10*cook/max(cook))
text(hatvalues(m1), rstudent(m1))
```

<br>

```{r} 
plot(hatvalues(m2), rstudent(m2), type = 'n')
#thresholds for residuals

abline(h = c(-2, 2), lty = 2)
#thresholds for leverages

abline(v = c(2,3)*3/56407, lty = 2)

cook <- sqrt(cooks.distance(m2))
points(hatvalues(m2), rstudent(m2), cex = 10*cook/max(cook))
text(hatvalues(m2), rstudent(m2))
```

<br>
<br>
<br>

#### Residuals distribution:

```{r}
qqPlot(m1)
```

<br>    

```{r}
qqPlot(m2) 
```

<br>
<br>
<br>

#### Tests for non-constant variances: 

```{r}
ncvTest(m1)
```

<br>

```{r}
library(lmtest)
bptest(m1, data = EVS, studentize = T)
```

There is a heteroscedasticity presence.

<br>

```{r}
ncvTest(m2)
```

<br>

```{r}
bptest(m2, data = EVS, studentize = T)
```

There is a heteroscedasticity in this model, too.

<br>

#### Residuals plots:

```{r}
library(plotmo)
plotres(m1, which = 1:4, caption = "m1")
```

<br>

```{r}
plotres(m2, which = 1:4, caption = "m2")
```

<br>
<br>
<br>
<br>
<br>

### Model 3

```{r}
m3 <- lm(surv ~ polit + gender + income + size_town + education + age, data = EVS)
tab_model(m3)
```

<br>
<br>
<br>

#### Check for multicollinearity:

```{r} 
vif(m3)  
```


<br>
<br>
<br>

#### Residuals distribution:

```{r}
qqPlot(m3) 
```

<br>

```{r}
ncvTest(m3)
```

<br>

```{r}
bptest(m3, data = EVS, studentize = T)
```

There is a heteroscedasticity presence.

<br>
<br>
<br>

#### Residuals plots:

```{r}
plotres(m3, which = 1:4, caption = "m3")
```

<br>
<br>
<br>
<br>
<br>

### Model 4

```{r}
m4 <- lm(surv ~ gender + income + size_town + education + age + polit*part, data = EVS)
tab_model(m4)
```

<br>
<br>
<br>

#### Check for multicollinearity:

```{r} 
vif(m4)
```

There is a multicollinearity presence for interaction effect.

<br>
<br>
<br>

#### Residuals distribution:

```{r}
qqPlot(m4) 
```

<br>

```{r}
ncvTest(m4)
```

<br>

```{r}
bptest(m4, data = EVS, studentize = T)
```

Based on non-constant variance test there is no heteroscedasticity,<br>
but studentized Breusch-Pagan test shows that there is a heteroscedasticity.

<br>
<br>
<br>

```{r}
plotres(m4, which = 1:4, caption = "m4")
```

<br>
<br>
<br>
<br>
<br>

### Model 5

```{r}
m5 <- lm(surv ~ polit + gender + income + size_town + education + age*part, data = EVS)
tab_model(m5)
```

<br>

```{r}  
interact_plot(m5, pred = age, modx = part, plot.points = FALSE) 
```


<br>
<br>
<br>

```{r}
vif(m5)
```

There is a multicollinearity presence for interaction effect.

<br>
<br>
<br>


#### Residuals distribution:

```{r}
qqPlot(m5) 
```

<br>

```{r}
ncvTest(m5) 
bptest(m5, data = EVS, studentize = T)
```

Based on non-constant variance test and studentized Breusch-Pagan test, there is a heteroscedasticity.

<br>
<br>
<br>

#### Residuals plots:

```{r}
plotres(m5, which = 1:4, caption = "m5")
```

<br>
<br>
<br>
<br>
<br>

### Non-constant error variance

(Heteroscedasticity)

```{r}  
spreadLevelPlot(m1)
```

<br>

```{r}  
spreadLevelPlot(m2)
```

<br>

```{r}  
spreadLevelPlot(m3)
```

<br>

```{r}  
spreadLevelPlot(m4)
```

<br>

```{r}  
spreadLevelPlot(m5)  
```


As there is a heteroscedasticity present, we need to use inference for estimated coefficients<br>
with a heteroscedasticity-consistent covariance matrix.

<br>
<br>
<br>
<br>
<br>

### Inference for Estimated Coefs

```{r}
library(sandwich)
coeftest(m1, vcovHC(m1, "HC1"))
```

<br>

```{r}
coeftest(m2, vcovHC(m2, "HC1"))
```

<br>

```{r}
coeftest(m3, vcovHC(m3, "HC1"))
```

<br>

```{r}
coeftest(m4, vcovHC(m4, "HC1"))
```

<br>

```{r}
coeftest(m5, vcovHC(m5, "HC1")) 
```

<br>
<br>
<br>

#### All three hypotheses stated here were confirmed.

<br>
<br>
<br>

#### The End. 


